Regression Cube Analysis of Cohort Study Data

### 2015-03-29 draft.tex
reworked Sylvias revision

% -Dependent & -Based in Überschriften groß, sicherheitshalber nochmal Steffi fragen
% oder: Dimension Reduction Depending on the Target Variable
\subsection{Target-Variable-Dependent Dimension Reduction}
% \com{Uli, Can you please proofread this?}
% %%
% The vast majority of features in an epidemiological data have no or a very low relation towards the target feature.
% %%
% Identifying these features und excluding them from the calculation can reduce the number of dimension significantly.
% %%
% The \emph{Correlation based Feature Selection} (CFS) algorithm is very popular in data mining for this task \cite{CFS}.
% %%
% It uses information entropy to select the features which have the most explanatory power towards the target feature.
% %%
% At the same time it tries to reduce the correlation between the independent features.
% %%
In epidemiological studies, manifold recordings lead to an abundance of features and thus a high-dimensional feature space.
%%
In general, many of them exhibit low or none correlation at all towards the target feature.
%%
Identifying irrelevant features and excluding them from the feature space considerably reduces computational costs and yields a comprehensible \emph{Regression Cube} representation.
%%
% nochmal checken wo $F$ eingeführt wird, hier oder im übernächsten Satz?
The correlation-based feature selection (CFS) \cite{CFS} aims to find a feature subset that maximizes the \emph{merit value} $M_F$ which is the ratio between the average feature-class and feature-feature dependencies of features in the feature set $F$.
%%
The dependency of a set of features utilizes the entropy-based information gain to measure the explanatory power towards the target feature.
%%
% Ggf. wurde $F$ schon oben eingeführt, TODO
Starting with an empty set of features $F$, the CFS algorithm iteratively adds the feature $f$ to $F$ that leads to the highest new merit-value $M_{F\cup f}$ and halts when no feature is left that would increase the merit.
%%
For example, if the \emph{body weight} has a strong explanatory power towards the target, it is likely that \emph{BMI} or \emph{waist circumference} exhibit similar correlations to the target.
%%
However, they strongly correlate with each other.
%%
The CFS algorithm will select the feature which has the largest explanatory power and discards the other dimensions.

@@

\paragraph{3D (Cube) View.}
%%
Introducing $Z$ creates a 3D \emph{Regression Cube} (Fig.~\ref{fig:Cube} b).
%%
$R^2$ values of each cube entry (\emph{voxel}) is mapped to opacity to reduce the overlap.
%%
The visualization of $R^2$ values derived from different regression cubes (e.g., $Z \sim X + Y$) is misleading, as they can be compared relatively, but not in precise numbers.
%%
% e.g. wenn man es wählen kann
% i.e. wenn es so ist
Therefore, the $R^2$ results of different regression methods are encoded using different colors (i.e., blue for linear regression and red for logistic regression).
%%
% Adverb immer vors Hauptverb oder Satzende
Thus, the cube can be easily extended using other regression types.
%%
For cubes with a fixed target feature, e.g., $Cancer \sim X + Y + Z$, no such encodings are required and the $z$ dimension can be compared directly.
%%
\\\\
Our goal is to create an overview visualization for a data set, but we also want to incorporate expert knowledge into the visualization by adapting the underlying formulas.
%%
These two approaches do not exclude each other, they rather underline the difference in purpose of the chosen formula.
%%
The different analysis approaches require different starting points using the \emph{Regression Cube}.

@@

% warum war da ein "either", sollte der Satz etwas anderes ausdrücken?
Hypotheses about the data are easily built up by relating dynamic variables with the regression operators.
%%
Furthermore, static features can be added for each regression formula.
%%
Here are a few examples:

>>>>

% -Dependent & -Based in Überschriften groß, sicherheitshalber nochmal Steffi fragen
% oder: Dimension Reduction Depending on the Target Variable
\subsection{Target-Variable-Dependent Dimension Reduction}
% \com{Uli, Can you please proofread this?}
% %%
% The vast majority of features in an epidemiological data have no or a very low relation towards the target feature.
% %%
% Identifying these features und excluding them from the calculation can reduce the number of dimension significantly.
% %%
% The \emph{Correlation based Feature Selection} (CFS) algorithm is very popular in data mining for this task \cite{CFS}.
% %%
% It uses information entropy to select the features which have the most explanatory power towards the target feature.
% %%
% At the same time it tries to reduce the correlation between the independent features.
% %%
In epidemiological studies, manifold recordings lead to an abundance of features and thus a high-dimensional feature space.
%%
In general, many of them exhibit low or none correlation at all towards the target feature.
%%
Identifying irrelevant features and excluding them from the feature space considerably reduces computational costs and yields a comprehensible \emph{Regression Cube} representation.
%%
The correlation-based feature selection (CFS) \cite{CFS} aims to find a feature subset that maximizes the \emph{merit value} $M_F$ which is the ratio between the average feature-class and feature-feature dependencies of features in a given feature set $F$.
%%
The dependency of a set of features utilizes the entropy-based information gain to measure the explanatory power towards the target feature.
%%
Starting with an empty set of features $F$, the CFS algorithm iteratively adds the feature $f$ to $F$ that leads to the highest new merit-value $M_{F\cup f}$ and halts when no feature is left that would increase the merit.
%%
For example, if the \emph{body weight} has a strong explanatory power towards the target, it is likely that \emph{BMI} or \emph{waist circumference} exhibit similar correlations to the target.
%%
However, they strongly correlate with each other.
%%
The CFS algorithm will select the feature which has the largest explanatory power and discards the other dimensions.

@@

\paragraph{3D (Cube) View.}
%%
Introducing $Z$ creates a 3D \emph{Regression Cube} (Fig.~\ref{fig:Cube} b).
%%
$R^2$ values of each cube entry (\emph{voxel}) is mapped to opacity to reduce the overlap.
%%
The visualization of $R^2$ values derived from different regression cubes (e.g., $Z \sim X + Y$) is misleading, as they can be compared relatively, but not in precise numbers.
%%
Therefore, the $R^2$ results of different regression methods are encoded using different colors (i.e., blue for linear regression and red for logistic regression).
%%
% Adverb immer vors Hauptverb oder Satzende
Thus, the cube can be easily extended using other regression types.
%%
For cubes with a fixed target feature, e.g., $Cancer \sim X + Y + Z$, no such encodings are required and the $z$ dimension can be compared directly.
%%
\\\\
Our goal is to create an overview visualization for a data set, but we also want to incorporate expert knowledge into the visualization by adapting the underlying formulas.
%%
These two approaches do not exclude each other, they rather underline the difference in purpose of the chosen formula.
%%
The different analysis approaches require different starting points using the \emph{Regression Cube}.

@@

Hypotheses about the data are easily built up by relating dynamic variables with the regression operators.
%%
Furthermore, static features can be added for each regression formula.
%%
Here are a few examples:
